---
title: "String substitution"
author: "Steve Simon"
date: "Wednesday, July 13, 2016"
output: html_document
---

here are a few simple examples of string substitution.

The sub function will look for a pattern, insert a replacement if it is found
for any string in a vector x.

```{r simple_example_with_sub}
example1 <- c("Steve", "Simon", "Thinker")
sub(pattern="e", replacement="3", x=example1)
```

Notice that the first "e" in the string "Steve" was changed to a "3" as was the "e"
in "Thinker". There was no match for "e" in "Simon", so it is left unchanged.

If you wanted to change more than one "e", you should use the gsub function.

```{r simple_example_with_gsub}
gsub(pattern="e", replacement="3", x=example1)
```

If you want to make a substitution of either a lower case or an upper case letter,
use the ignore.case=TRUE option.

```{r ignoring_case}
sub(pattern="s", replacement="3", x=example1, ignore.case=TRUE)
```

By default, the sub and gsub functions use regular expressions. Regular 
expressions are very powerful, but they require special treatment for
particular symbols such as the period and the asterisk. So searching,
for example, for the string like "-0.2" has the potential for trouble.

If you are just starting out with string substitutions, you may want to avoid
regular expressions until you get more comfortable with the whole process.
Just use the option fixed=TRUE to disable regular expressions.

```{r turninging_off_regular_expressions}
example2 <- c("0.11", "0.20", "0.29")
gsub(pattern=".", replacement=",", x=example2, fixed=TRUE)
```

Try this function without fixed=TRUE and see what happens.

```{r what_if}
gsub(pattern=".", replacement=",", x=example2)
```

The alternative, if you want to stay with regular expressions, is to
place a double backslash in front of most special characters.

```{r double_backslash}
gsub(pattern="\\.", replacement=",", x=example2)
```

Here are a few simple tricks that regular expressions can do. The caret symbol (^)
will find a match only if it is at the start of a your string and the dollar symbol ($)
will find a match only if it is at the end of your string.

```{r start_and_end}
sub(pattern="^0", replacement="X", x=example2)
sub(pattern="0$", replacement="X", x=example2)
```

You can use square brackets to match more than one thing.

```{r square_brackets}
gsub(pattern="[012]", replacement="X", x=example2)
```
