---
title: "String substitution"
author: "Steve Simon"
date: "Wednesday, July 13, 2016"
output: html_document
---

here are a few simple examples of string substitution.

The sub function will look for a pattern, insert a replacement if it is found
for any string in a vector x.

```{r simple_example_with_sub}
example_strings <- c("Steve", "Simon", "Thinker")
sub(pattern="e", replacement="3", x=example_strings)
```

Notice that the first "e" in the string "Steve" was changed to a "3" as was the "e"
in "Thinker". There was no match for "e" in "Simon", so it is left unchanged.

If you wanted to change more than one "e", you should use the gsub function.

```{r simple_example_with_gsub}
gsub(pattern="e", replacement="3", x=example_strings)
```

By default, the sub and gsub functions use regular expressions. Regular 
expressions are very powerful, but they require special treatment for
particular symbols such as the period and the asterisk. So searching,
for example, for the string "0.2" is not that easy.

If you are just starting out with string substitutions, you may want to avoid
regular expressions until you get more comfortable with the whole process.

```{r turninging_off_regular_expressions}
example_strings <- c("0.11", "0.22", "0.33")
sub(pattern="0.2", replacement=".2", x=example_strings, fixed=TRUE)
